---
title: "Quick Plots for Toxics Report"
author: "Ben Best"
date: "January 10, 2018"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
```

# 2018-10-12

Can I ask you to create a basic graph with the attached lead data w/ indicators for three numeric concentrations (CTR 0.0025 mg/L, Benchmark 0.0816 mg/L & NAL 0.26 mg/L). Its fine if some of the highest concentrations actually go off the graph.  I'm imagining the horizontal axis is years/dates and the vertical is lead concentrations.  


# 1420.2 Sample History

## Timeseries of Pb per Facility over Years

```{r pb-facilities, fig.cap="Figure #. Timeseries of lead values per facility over years. Median values (point) are derived from the average of the high and low values (error bars) extracted from the 1420.2 reports. Limits (dashed lines) are shown for EPA Benchmark, NAL and CTR."}
library(tidyverse)
library(readxl)
library(here)
library(DT)

d_xlsx <- here("data/1420.2 Sample History.xlsx")

d <- read_excel(d_xlsx) %>%
  mutate(
    yr_beg = `AR YEARS` %>%
      str_replace("(.*)-(.*)", "\\1") %>%
      as.integer())

d2 = d %>%
  select(facility=FACILITY, yr_beg, pb_hi=`Pb (high)`, pb_lo=`Pb (low)`) %>%
  filter(!is.na(pb_hi) & !is.na(pb_lo)) %>%
  rowwise() %>%
  mutate(
    pb_md = mean(c(pb_hi, pb_lo))) #View(d2)

p_pb_facilities <- ggplot(data=d2, aes(
  x = yr_beg, 
  y = pb_md, ymin = pb_lo, ymax = pb_hi,
  color =facility)) +
  geom_errorbar(width=0.2) + geom_point() + geom_line()

p_pb_limits <- function(p){
  p +   
    # add hlines alphabetically
    # geom_hline(aes(yintercept=0.0816, linetype="Benchmark"), color = "blue") +
    # geom_hline(aes(yintercept=0.0025, linetype="CTR"), color = "green") +
    # geom_hline(aes(yintercept=0.262 , linetype="NAL"), color = "red") +
    geom_hline(aes(yintercept=0.0816, linetype="Benchmark")) +
    geom_hline(aes(yintercept=0.0025, linetype="CTR")) +
    geom_hline(aes(yintercept=0.262 , linetype="NAL")) +
    scale_linetype_manual(
      name = "Limits", values = c(2, 3, 4)) #, 
      #guide = guide_legend(override.aes = list(color = c("blue", "green", "red"))))
}

p_pb_labs <- function(p, sub="Lead (log10 transformed) by year per facility"){
  p +   labs(subtitle=sub, 
       y="Pb (mg/L)", 
       x="Year", 
       title="Lead", 
       caption = "Source: 1420.2 Reports")
}

p_pb_facilities %>% 
  p_pb_limits() %>%
  p_pb_labs("Lead (log10 transformed) by year per facility")
```

## Timeseries of Pb per Facility over Years, log10

```{r pb-facilities-log10, fig.cap="Figure #. Timeseries of lead values per facility over years. Median values (point) are derived from the average of the high and low values (error bars) extracted from the 1420.2 reports. Note the log10 transformation of the y axis applied to show facility exceedance around the limits (dashed lines) for EPA Benchmark, NAL and CTR."}
p_pb_facilities %>% 
  p_pb_limits() %>%
  p_pb_labs("Lead by year per facility") + 
  scale_y_log10(breaks=10^(-2:2))
```

## Boxplot of Pb across Facilities over Years

```{r pb-box, fig.cap="Figure #. Boxplot of lead values across facilities over years. Values for the boxplot are generated by including the high, low and average of the high and low values extracted from the 1420.2 reports. Limits (dashed lines) are shown for EPA Benchmark, NAL and CTR."}
d3 <- d2 %>%
  gather(key, val, -facility, -yr_beg)

p_pb_box <- ggplot(data=d3, aes(x=as.factor(yr_beg), y=val)) +
  geom_boxplot() 

p_pb_box %>% 
  p_pb_limits() %>%
  p_pb_labs("Lead (log10 transformed) by year across facilities")
```

## Boxplot of Pb across Facilities over Years, log10

```{r pb-box-log10, fig.cap="Figure #. Boxplot of lead values across facilities over years. Values for the boxplot are generated by including the high, low and average of the high and low values extracted from the 1420.2 reports. Note the log10 transformation of the y axis applied to show facility exceedance around the limits (dashed lines) for EPA Benchmark, NAL and CTR."}
p_pb_box %>% 
  p_pb_limits() %>%
  p_pb_labs("Lead (log10 transformed) by year across facilities") + 
  scale_y_log10(breaks=10^(-2:2))
```


# Trojan 1994 to Present 

## Bar Plot

```{r bar}
d_xlsx <- here("data/Trojan 1994 to Present.xlsx")
d <- read_excel(d_xlsx, na="no data")

d_na <- d %>% 
  filter(is.na(`Pb Concentration (mg/L)`)) %>%
  mutate(
    lbl = "NA",
    y   = 0)

g <- ggplot(d, aes(x=Year, y=`Pb Concentration (mg/L)`))
g + geom_col() +
  geom_text(
    aes(x=Year, y=y, label=lbl), data=d_na,
    size=3, vjust=-0.5)
```

## Point Plot

```{r point}
g + geom_point()
```

## Point & Line Plot

```{r point-line}
g + geom_point() + geom_line()
```

## Data

```{r tbl}
datatable(d)
```

